Inference using context-dependent models of mutation

### 2021-10-11 writeups/context-dependence.tex
Quick pass through the edits

\begin{example}[Genomic GC content]
    A sequence of genome can be written using A, C, G, and T;
    in the most general model of independent mutation across sites, each of the 12 possible transitions occurs at its own rate.
    Furthermore, it is a well-known observation that in many species,
    adjacent, methylated CG dinucleotides (``CpG sites'') have a much higher mutation rate to TG and CA
    than either single nucleotide change under the independent model.
    Embellishing the single-nucleotide model with this additional rate results in the model defined by
    \begin{center}
      \begin{tabular}{c@{\quad$\to$\quad}c@{\quad at rate\quad }cc}
        $x$  &  $y$  &  $m_{xy}$ & when $x \neq y \in \{\nA,\nC,\nG,\nT\}$  \\
        \nC\nG   &  \nT\nG   &  $\gamma$ & \\
        \nC\nG   &  \nC\nA   &  $\gamma$ &
      \end{tabular} ,
    \end{center}
    where $\gamma$ is the additional CpG rate above the base mutation rate.
    Note that for a given pair of sequences,
    a change $\nA\nC\nG \to \nA\nT\nG$ could have occurred by a single $\nC \to \nT$ mutation,
    or by a $\nC\nG \to \nT\nG$ mutation;
    in such cases
    the total instantaneous rate at which ACG changes to ATG is the sum of the rates
    (in this example, $\gamma + m_{\nC\nT}$).
    % Although this means there are many possible parameterizations,
    % note that the model is identifiable, not overparameterized.

>>>>

\begin{example}[Genomic GC content]
    A sequence of genome can be written using A, C, G, and T;
    in the most general model of independent mutation across sites, each of the 12 possible transitions occurs at its own rate.
    Furthermore, it is a well-known observation that in many species,
    adjacent, methylated CG dinucleotides (``CpG sites'') have a much higher mutation rate to TG and CA
    than either single nucleotide change under the independent model.
    Embellishing the single-nucleotide model with this additional rate results in the model defined by
    \begin{center}
      \begin{tabular}{c@{\quad$\to$\quad}c@{\quad at rate\quad }cc}
        $x$  &  $y$  &  $m_{xy}$ & when $x \neq y \in \{\nA,\nC,\nG,\nT\}$  \\
        \nC\nG   &  \nT\nG   &  $\gamma$ & \\
        \nC\nG   &  \nC\nA   &  $\gamma$ &
      \end{tabular} ,
    \end{center}
    where $\gamma$ is the additional CpG rate above the base mutation rate.
    Note that for a given pair of sequences,
    a change $\nA\nC\nG \to \nA\nT\nG$ could have occurred by a $\nC \to \nT$ mutation
    via the independent-sites model,
    or by a $\nC\nG \to \nT\nG$ mutation;
    in such cases
    the total instantaneous rate at which ACG changes to ATG is the sum of the rates
    (in this example, $\gamma + m_{\nC\nT}$).
    % Although this means there are many possible parameterizations,
    % note that the model is identifiable, not overparameterized.
